Ship-based lidar evaluation of Southern Ocean clouds in the storm-resolving general circulation model ICON, and the ERA5 and MERRA-2 reanalyses

### 2024-12-09 manuscript_agu.tex
AGU template

% Related publications:
%
% Takasuka et al. (2024), A protocol and analysis of year-long simulations of global storm-resolving models and beyond
% Knight et al. (2024), Cloud properties and boundary layer stability above Southern Ocean sea ice and coastal Antarctica
% Zheng et al. (2024), Using Satellite and ARM Observations to Evaluate Cold Air Outbreak Cloud Transitions in E3SM Global Storm-Resolving Simulations
% Schmidt et al. (2024), Effects of vertical grid spacing on the climate simulated in the ICON-Sapphire global storm-resolving model
% Fons et al. (2024), Investigating the sign of stratocumulus adjustments to aerosols in the global storm-resolving model ICON
% Fiddes et al. (2024), A machine learning approach for evaluating Southern Ocean cloud radiative biases in a global atmosphere model
% Mooers et al. (2023), Comparing storm resolving models and climates via unsupervised machine learning
% Zhang et al. (2023), Understanding models' global sea surface temperature bias in mean state: from CMIP5 to CMIP6
% Grundner (2023), Data-Driven Cloud Cover Parameterizations for the ICON Earth System Model Using Deep Learning and Symbolic Regression [thesis]
% Luo et al. (2023), Origins of Southern Ocean warm sea surface temperature bias in CMIP6 models
% Sherriff-Tadano et al. (2023), Southern Ocean Surface Temperatures and Cloud Biases in Climate Models Connected to the Representation of Glacial Deep Ocean Circulation
% Danker et al. (2022), Exploring relations between cloud morphology, cloud phase, and cloud radiative properties in Southern Ocean's stratocumulus clouds
% Possner et al. (2022), Resolution Dependence of Southern Ocean Mixed-Phase Clouds in ICON
% Mauritsen et al. (2022), Early Development and Tuning of a Global Coupled Cloud Resolving Model, and its Fast Response to Increasing CO2
% Wall et al. (2022), Observational Constraints on Southern Ocean Cloud-Phase Feedback
% Grundner et al. (2022), Deep learning based cloud cover parameterization for ICON
% Cesana et al. (2022), Southern Ocean Solar Reflection Biases in CMIP6 Models Linked to Cloud Phase and Vertical Structure Representations
% Fiddes et al. (2022), Southern Ocean cloud and shortwave radiation biases in a nudged climate model simulation: does the model ever get it right?
% Zhao et al. (2022), Compensating Errors in Cloud Radiative and Physical Properties over the Southern Ocean in the CMIP6 Climate Models
% Konsta et al. (2022), Low-Level Marine Tropical Clouds in Six CMIP6 Models Are Too Few, Too Bright but Also Too Compact and Too Homogeneous
% Lin and Yu (2022), The potential impact of model horizontal resolution on the simulation of atmospheric cloud radiative effect in CMIP6 models
% Ramadoss et al. (2021), An evaluation of kilometre scale ICON simulations of mixed-phase stratocumulus over the Southern Ocean durin CAPRICORN [poster]
% Miao et al. (2021), A Regime-based Investigation into the Errors of CMIP6 Simulated Cloud Radiative Effects Using Satellite Observations
% Roh et al. (2021), Intercomparison of cloud properties in DYAMOND simulations over the Atlantic Ocean
% Caldwell et al. (2021), Convection-Permitting Simulations With the E3SM Global Atmosphere Model
% Schuddeboom and McDonald (2021), The Southern Ocean Radiative Bias, Cloud Compensating Errors, and Equilibrium Climate Sensitivity in CMIP6 Models
% Ramadoss et al. (2020), Simulating mixed-phase cloud properties with ICON around the CAPRICORN field campaign at the kilometre scale [poster]
% Heim et al. (2021), Inter-model Variability in Convection-Resolving Simulations of Subtropical Marine Low Clouds
% Seiki and Roh (2020), Improvements in Supercooled Liquid Water Simulations of Low-Level Mixed-Phase Clouds over the Southern Ocean Using a Single-Column Model
% Stevens et al. (2020), The added value of large-eddy and storm-resolving models for simulating clouds and precipitation
% Atlas et al. (2020), How well do large‐eddy simulations and global climate models represent observed boundary layer structures and low clouds over the summertime Southern Ocean?
% Korn et al. (2020), ICON‐O: The Ocean Component of the ICON Earth System Model—Global Simulation Characteristics and Local Telescoping Capability
% Vignesh et al. (2020), Assessment of CMIP6 Cloud Fraction and Comparison with Satellite Observations
% Chen et al. (2022), Evaluation of Simulated Cloud Diurnal Variation in CMIP6 Climate Models
% Jian et al. (2020), Evaluation of the CMIP6 planetary albedo climatology using satellite observations
% Satoh et al. (2019), Global Cloud-Resolving Models
% Hyder et al. (2018), Critical Southern Ocean climate model biases traced to atmospheric model cloud errors

\documentclass[12pt,a4paper]{article}
\usepackage{fontspec}
\usepackage{microtype}
\usepackage[margin=2cm]{geometry}
\usepackage[citecolor=black]{hyperref}
\usepackage[authoryear,round]{natbib}
\usepackage{doi}
\usepackage{titlesec}
\usepackage{caption}
\DeclareCaptionLabelSeparator{pipe}{ | }
\captionsetup{labelfont={bf,sf},labelsep=pipe}
\usepackage{graphicx}
\usepackage{lineno}
\linenumbers
\setsansfont{Source Sans Pro}
\setmainfont{EB Garamond}
\setmonofont[Scale=0.8]{SpaceMono}[
	Path=./fonts/,
	Extension=.ttf,
	UprightFont=*-Regular,
	BoldFont=*-Bold,
	ItalicFont=*-Italic,
	BoldItalicFont=*-BoldItalic
]
\hypersetup{
	colorlinks=true,
	urlcolor=blue,
	pdftitle={Ship-based lidar evaluation of Southern Ocean clouds in the storm-resolving general circulation model ICON, and the ERA5 and MERRA-2 reanalyses},
	pdfauthor={Peter Kuma, Frida A.-M. Bender, Adrian J. McDonald, Simon P. Alexander, Greg M. McFarquhar, John J. Cassano, Graeme E. Plank, Sean Hartery, Simon Parsons, Sally Garrett, and Alex J. Schuddeboom}
}
\bibliographystyle{abbrvnat}
\titleformat*{\section}{\large\bfseries\sffamily}
\titleformat*{\subsection}{\normalsize\bfseries\sffamily}
\titlespacing*{\section}{0pt}{14pt}{6pt}
\titlespacing*{\subsection}{0pt}{13pt}{4pt}
\parskip=6pt
\parindent=0pt

@@

\fontsize{13pt}{15pt}\selectfont

\begin{center} \Large \sffamily\textbf{Ship-based lidar evaluation of Southern Ocean clouds in the storm-resolving general circulation model ICON and the ERA5 and MERRA-2 reanalyses}\\[0.4cm]
\normalsize Peter Kuma$^\mathsf{1,2,\star}$, Frida A.-M. Bender$^\mathsf{1,2}$, Adrian J. McDonald$^\mathsf{3}$, Simon P. Alexander$^\mathsf{4,5}$, Greg M. McFarquhar$^\mathsf{6,7}$, John J. Cassano$^\mathsf{8,9,10}$, Graeme E. Plank$^\mathsf{3}$, Sean Hartery$^\mathsf{11}$, Simon Parsons$^\mathsf{12}$, Sally Garrett$^\mathsf{13}$, and Alex J. Schuddeboom$^\mathsf{3}$\\[0.4cm]
\footnotesize
$^\mathsf{1}$Department of Meteorology (MISU), Stockholm University, Stockholm, Sweden\\
$^\mathsf{2}$Bolin Centre for Climate Research, Stockholm University, Stockholm, Sweden\\
$^\mathsf{3}$School of Physical and Chemical Sciences, University of Canterbury, Christchurch, Aotearoa/New Zealand\\
$^\mathsf{4}$Australian Antarctic Division, Kingston, Tasmania, Australia\\
$^\mathsf{5}$Australian Antarctic Program Partnership, Institute for Marine and Antarctic Studies, University of Tasmania, Hobart, Tasmania, Australia\\
$^\mathsf{6}$Cooperative Institute of Severe and High Impact Weather Research and Operations, University of Oklahoma, Norman, OK, USA\\
$^\mathsf{7}$School of Meteorology, University of Oklahoma, Norman, OK, USA\\
$^\mathsf{8}$Cooperative Institute for Research in Environmental Sciences, University of Colorado, Boulder, CO, USA\\
$^\mathsf{9}$National Snow and Ice Data Center, University of Colorado, Boulder, CO, USA\\
$^\mathsf{10}$Department of Atmospheric and Oceanic Sciences, University of Colorado, Boulder, CO, USA\\[0.2cm]
$^\mathsf{11}$Department of Physics \& Atmospheric Science, Dalhousie University, Halifax, Canada\\
$^\mathsf{12}$New South Wales Department of Planning and Environment, Sydney, New South Wales, Australia\\
$^\mathsf{13}$New Zealand Defence Force, Wellington, New Zealand\\
$^\star$Corresponding author: Peter Kuma (\href{mailto:peter.kuma@misu.su.se}{peter.kuma@misu.su.se})\\[0.4cm]
\normalsize \today\\[0.4cm]

\end{center}

\section*{Abstract}

Global storm-resolving models (GSRMs) are the next avenue of climate modelling.
Among them is a 5-km version of the Icosahedral Nonhydrostatic Weather and
Climate Model (ICON). The high resolution means that parameterizations of
convection and clouds including subgrid-scale clouds are forgone, relying on
explicit their simulation, but still using microphysics and turbulence
parametrization.  Standard-resolution (10--100-km) models, which use
subgrid-scale convection and cloud parametrization, have substantial cloud
biases over the Southern Ocean (SO), affecting radiation and sea surface
temperature.  We evaluated SO clouds in ICON and the ERA5 and MERRA-2
reanalyses.  The SO is dominated by low clouds, which cannot be observed
accurately from space due to overlapping clouds, attenuation, and ground
clutter.  Instead, we analysed about 2400 days of lidar observations from 31
voyages and a station during 2010--2021 using a ground-based lidar simulator.
ICON and the reanalyses underestimate the total cloud fraction by about 10 and
20\%, respectively. ICON and ERA5 overestimate the cloud occurrence peak at
about 500 m, potentially explained by their lifting condensation levels being
too high.  The reanalyses strongly underestimate near-surface clouds or fog.
MERRA-2 tends to underestimate cloud occurrence at all heights.  In daily cloud
cover, ICON and the reanalyses tend to be about 1 and 2 oktas clearer,
respectively.  Conditions of weak stability are associated with the largest
errors in ICON and the reanalyses.  Compared to radiosondes, potential
temperature is accurate in the reanalyses to about 0.5 K, but ICON
underestimates stability over the low-latitude SO and too humid in the boundary
layer. MERRA-2 is too humid at all heights. SO cloud biases remain a
substantial issue in the GSRM, but are an improvement over the lower-resolution
reanalyses. Explicitly resolved convection and cloud processes (except
microphysics) were not enough to address the model cloud biases.

@@

Increasing climate model resolution is one way of improving model accuracy of
representation of the climate system \citep{mauritsen2022}. It has been
practiced since the advent of climate modelling as more computational power,
memory, and storage capacity become available. It is, however, often not as
easy as changing the grid size because of the complex interplay between model
dynamics and physics, which necessitates adjusting and tuning all components
together.  Increasing resolution is of course limited by the available
computational power and a trade-off with increasing parameterization
complexity, which is another way of improving model accuracy.  Current
computational availability and acceleration from general-purpose computing on
graphics processing units (GPUs) is progressing to enable km-scale (also called
k-scale) Earth system models (ESMs) and coupled atmosphere--ocean general
circulation models (AOGCMs) in research conditions today and operationally in
the forthcoming years.  Therefore, it represents a natural advance in climate
modelling.  Global storm-resolving models (GSRMs) are emerging as a new front
in the development of high-resolution global climate models, with horizontal
grid resolutions of about 2--8 km \citep{satoh2019,stevens2019}. This is enough
to resolve mesoscale convective storms, but smaller-scale convective plumes and
cloud structure remain unresolved. At an approximately 5-km scale,
non-hydrostatic processes also become important \citep{weisman1997}, and for
this reason such models are generally non-hydrostatic. The terms global
cloud-resolving models or global convection-permitting/-resolving models are
also sometimes used interchangeably with GSRMs but imply that clouds or
convection are resolved explicitly, which is not entirely true for GSRMs, as
this would require an even higher horizontal resolution \citep{satoh2019}.
Representative of these efforts is the DYnamics of the Atmospheric general
circulation Modeled On Non-hydrostatic Domains (DYAMOND) project
\citep{stevens2019,dyamond}, which is an intercomparison of nine global GSRMs
over two 40-day time periods in summer (1 August -- 10 September 2016) and
winter (20 January -- 1 March 2020). A new one-year GSRM intercomparison is
currently proposed by \cite{takasuka2024}, with the hope of also evaluating the
seasonal cycle and large-scale circulation.  An alternative to using a
computationally costly GSRM is to train an artificial neural network on GSRM
output and use it for subgrid-scale clouds, as done with the GSRM ICON by
\cite{grundner2022} and \cite{grundner2023}.

nextGEMS is a project \citep{nextgems} focused on the research and development
of GSRMs at multiple modelling centres and universities in Europe.  The project
also develops GSRM versions of the Icosahedral Nonhydrostatic Weather and
Climate Model (ICON; \cite{hohenegger2023}), the Integrated Forecasting System
[IFS; \cite{ifs2023}], and their ocean components at eddy-resolving
resolutions: ICON-O \citep{korn2022} coupled with ICON and
Finite-Element/volumE Sea ice-Ocean Model [FESOM; \cite{wang2014}] and Nucleus
for European Modelling of the Ocean [NEMO; \cite{madec2023}] coupled with IFS.
The project has so far produced ICON and IFS simulations with three development
versions called Cycle 1--3 and a pre-final version, with a final production
version planned by the end of the project. nextGEMS is not the only project
developing GSRMs; other GSRMs (or GSRM versions of climate models) currently in
development include: Convection-Permitting Simulations With the E3SM Global
Atmosphere Model [SCREAM; \cite{caldwell2021}], Atmospheric Model [NICAM;
\cite{satoh2008}], Unified Model (UM), eXperimental System for High-resolution
modeling for Earth-to-Local Domain [X-SHiELD; \cite{shield}], Action de
Recherche Petite Echelle Grande Echelle-NonHydrostatic version [ARPEGE-NH;
\cite{bubnova1995,voldoire2017}], Finite-Volume Dynamical Core on the Cubed
Sphere [FV3, \cite{lin2004}], the National Aeronautics and Space Administration
(NASA) Goddard Earth Observing System global atmospheric model version 5
[GEOS5; \cite{putman2011}], Model for Prediction Across Scales [MPAS;
\cite{skamarock2012}], and System for Atmospheric Modeling [SAM;
\cite{khairoutdinov2003}].

Multiple cloud properties have an effect on shortwave (SW) and longwave (LW)
radiation. To first order, the total cloud fraction, cloud phase, and the
liquid and ice water path are the most important cloud properties influencing
SW and LW radiation. These properties are  in turn influenced by the
atmospheric thermodynamics, convection and circulation, and both the indirect
and direct effects of aerosols. Second order effects on SW and LW radiation are
associated with the cloud droplet size distribution, ice crystal habit, cloud
lifetime, and direct radiative interaction with aerosols.  In the
6$^\mathrm{th}$ phase of the Coupled Model Intercomparison Project [CMIP6;
\cite{eyring2016}], the cloud feedback has increased relative to CMIP5
\citep{zelinka2020}, which is one of the main reasons for the higher climate
sensitivity of CMIP6 models.

The Southern Ocean (SO) is known to be a problematic region for climate model
biases \citep{schuddeboom2021,hyder2018,cesana2022,zhao2022} due to a lack of
surface and in situ observations and being a lower priority region for
numerical weather prediction (NWP) and climate model development because of its
distance from populated areas.  Nevertheless, radiation biases and changes over
an area of its size have a substantial influence on the global climate
\citep{rintoul2011}, such as affecting the Earth radiation balance, ocean heat
and carbon uptake \citep{williams2023}, and the SO is an important part of the
global ocean conveyor belt \citep{wang2014}. In general, marine clouds have a
disproportionate effect on top of atmosphere (TOA) SW radiation due to the
relatively low albedo of the sea surface.  The relative longitudinal symmetry
of the SO means that model cloud biases tend to be similar across longitudes.

Here, we conventionally refer to the SO as ocean regions south of 40°S,
low-latitude SO as 40--55°S and high-latitude SO as south of 55°S. The reason
for this dividing latitude is to split the SO into about two equal zones, as
well as the results by \cite{schuddeboom2021} (Fig. 2b), showing a contrast in
CMIP model radiation biases, \cite{schuddeboom2019} (Fig. 2) and
\cite{kuma2020} (Fig. 3) showing contrasting radiation biases in the Hadley
Centre Global Environmental Model, and \cite{cesana2022} showing contrasting
cloud biases due to the 0°C isotherm reaching the surface at 55°S. The findings
of \cite{niu2024}, however, support a different dividing line of 62°S based on
cloud condensation nuclei concentration.

SO radiation biases have been relatively large and systematic compared to the
rest of the globe since at least CMIP3 \citep{trenberth2010}, and the SO SW
cloud radiative effect (CRE) bias is still positive in eight analysed CMIP6
models analysed by \cite{schuddeboom2021} over the high-latitude SO, whereas
over the low-latitude SO it tends to be more neutral or negative in some
models. Too much absorbed SW radiation over the SO was also identified in the
GSRM SCREAM \cite{caldwell2021}. Compensating biases are possible, such as the
`too few too bright' cloud bias, characterised by too small cloud fraction and
too large cloud albedo \citep{wall2017,kuma2020}, previously described by
\cite{webb2001}, \cite{weare2004}, \cite{zhang2005}, \cite{karlsson2008},
\cite{nam2012}, \cite{klein2013}, and \cite{bender2017} in other regions and
models, which means that a model can maintain a reasonable SW radiation balance
by reflecting too much SW radiation from clouds, but these cover too small
area.  A study by \cite{konsta2022} showed that this type of bias is still
present in six analysed CMIP6 models in tropical marine clouds, using the
GCM-Oriented Cloud-Aerosol Lidar and Infrared Pathfinder Satellite Observation
(CALIPSO) Cloud Product [CALIPSO--GOCCP; \cite{chepfer2010}] and Polarization
\& Anisotropy of Reflectances for Atmospheric Sciences coupled with
Observations from a Lidar [PARASOL; \cite{lier2008}] as a reference. They
suggest improper simulation of subgrid-scale cloud heterogeneity as a cause.
Compensating cloud biases in the Australian Community Climate and Earth System
Simulator (ACCESS) – Atmosphere-only model version 2 (AM2) over the SO were
analysed by \cite{fiddes2022} and \cite{fiddes2024}.  \cite{possner2022} showed
that over the SO, the DYAMOND GSRM ICON underestimates low-level cloud fraction
on the order of 30\% [relative to Moderate Resolution Imaging Spectroradiometer
(MODIS) data] and overestimates downwelling TOA SW radiation by approximately
10 Wm$^\mathrm{-2}$ [relative to the Clouds and the Earth’s Radiant Energy
System (CERES) data] in the highest model resolution run (2.5 km).
\cite{zhao2022} reported a similar SW radiation bias in five analysed CMIP6
models over high-latitude SO and the total cloud fraction underestimation on
the order of 10\% over the entire 40--60°S SO. Recently, \cite{ramadoss2024}
analysed 48 hours of km-scale ICON limited area model NWP simulations over a SO
region adjacent to Tasmania against the Clouds, Aerosols, Precipitation,
Radiation, and atmospherIc Composition Over the southeRn oceaN (CAPRICORN)
voyage cloud and precipitation observations \cite{mcfarquhar2021}. They found
the ICON cloud optical thickness was underestimated relative to Himawari‐8
satellite observations, but also identified large differences in cloud top
phase.

In general, sea surface temperature (SST) biases in the SO can originate either
in the atmosphere \citep{hyder2018}, caused by too much shortwave heating of
the surface, too little longwave cooling of the surface, or in the ocean
circulation.  Interactions of both are also possible, for example SST affecting
clouds and clouds affecting the surface radiation. Using the European Centre
for Medium-Range Weather Forecasts (ECMWF) Reanalysis 5 (ERA5) as a reference,
\cite{zhang2023} have shown that SST biases have improved in CMIP6 compared to
CMIP5, with SST overall increasing in the later CMIP phase. However, over the
SO this resulted in an even higher positive bias, especially in the Atlantic
Ocean (AO) sector of the SO, increasing by up to 1°C.  \cite{luo2023}
identified that the SO SST bias in an ensemble of 18 CMIP6 models originates
not from the surface heat and radiation fluxes (using reanalyses as a
reference), but from a warm bias in the Northern Atlantic Deep Water.

>>>>

\documentclass[draft]{agujournal2019}
\usepackage{url} %this package should fix any errors with URLs in refs.
\usepackage{lineno}
%\usepackage[inline]{trackchanges} %for better track changes. finalnew option will compile document with changes incorporated.
\usepackage{soul}
\linenumbers

\draftfalse

\journalname{JGR: Atmospheres}

@@

\title{Ship-based lidar evaluation of Southern Ocean clouds in the storm-resolving general circulation model ICON, and the ERA5 and MERRA-2 reanalyses}

\authors{Peter Kuma\affil{1,2}\thanks{}, Frida A.-M. Bender\affil{1,2}, Adrian J. McDonald\affil{3}, Simon P. Alexander\affil{4,5}, Greg M. McFarquhar\affil{6,7}, John J. Cassano\affil{8,9,10}, Graeme E. Plank\affil{3}, Sean Hartery\affil{11}, Simon Parsons\affil{12}, Sally Garrett\affil{13}, and Alex J. Schuddeboom\affil{3}}

\affiliation{1}{Department of Meteorology (MISU), Stockholm University, Stockholm, Sweden}
\affiliation{2}{Bolin Centre for Climate Research, Stockholm University, Stockholm, Sweden}
\affiliation{3}{School of Physical and Chemical Sciences, University of Canterbury, Christchurch, Aotearoa/New Zealand}
\affiliation{4}{Australian Antarctic Division, Kingston, Tasmania, Australia}
\affiliation{5}{Australian Antarctic Program Partnership, Institute for Marine and Antarctic Studies, University of Tasmania, Hobart, Tasmania, Australia}
\affiliation{6}{Cooperative Institute of Severe and High Impact Weather Research and Operations, University of Oklahoma, Norman, OK, USA}
\affiliation{7}{School of Meteorology, University of Oklahoma, Norman, OK, USA}
\affiliation{8}{Cooperative Institute for Research in Environmental Sciences, University of Colorado, Boulder, CO, USA}
\affiliation{9}{National Snow and Ice Data Center, University of Colorado, Boulder, CO, USA}
\affiliation{10}{Department of Atmospheric and Oceanic Sciences, University of Colorado, Boulder, CO, USA}
\affiliation{11}{Department of Physics \& Atmospheric Science, Dalhousie University, Halifax, Canada}
\affiliation{12}{New South Wales Department of Planning and Environment, Sydney, New South Wales, Australia}
\affiliation{13}{New Zealand Defence Force, Wellington, Aotearoa/New Zealand}

\correspondingauthor{Peter Kuma}{peter@peterkuma.net}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% KEY POINTS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%  List up to three key points (at least one is required)
%  Key Points summarize the main points and conclusions of the article
%  Each must be 140 characters or fewer with no special characters or punctuation and must be complete sentences

% Example:
% \begin{keypoints}
% \item	List up to three key points (at least one is required)
% \item	Key Points summarize the main points and conclusions of the article
% \item	Each must be 140 characters or fewer with no special characters or punctuation and must be complete sentences
% \end{keypoints}

\begin{keypoints}
\item enter point 1 here
\item enter point 2 here
\item enter point 3 here
\end{keypoints}

\begin{abstract}
Global storm-resolving models (GSRMs) are the next avenue of climate modelling.
Among them is a 5-km version of the Icosahedral Nonhydrostatic Weather and
Climate Model (ICON). The high resolution means that parameterizations of
convection and clouds including subgrid-scale clouds are forgone, relying on
explicit their simulation, but still using microphysics and turbulence
parametrization.  Standard-resolution (10--100-km) models, which use
subgrid-scale convection and cloud parametrization, have substantial cloud
biases over the Southern Ocean (SO), affecting radiation and sea surface
temperature.  We evaluated SO clouds in ICON and the ERA5 and MERRA-2
reanalyses.  The SO is dominated by low clouds, which cannot be observed
accurately from space due to overlapping clouds, attenuation, and ground
clutter.  Instead, we analysed about 2400 days of lidar observations from 31
voyages and a station during 2010--2021 using a ground-based lidar simulator.
ICON and the reanalyses underestimate the total cloud fraction by about 10 and
20\%, respectively. ICON and ERA5 overestimate the cloud occurrence peak at
about 500 m, potentially explained by their lifting condensation levels being
too high.  The reanalyses strongly underestimate near-surface clouds or fog.
MERRA-2 tends to underestimate cloud occurrence at all heights.  In daily cloud
cover, ICON and the reanalyses tend to be about 1 and 2 oktas clearer,
respectively.  Conditions of weak stability are associated with the largest
errors in ICON and the reanalyses.  Compared to radiosondes, potential
temperature is accurate in the reanalyses to about 0.5 K, but ICON
underestimates stability over the low-latitude SO and too humid in the boundary
layer. MERRA-2 is too humid at all heights. SO cloud biases remain a
substantial issue in the GSRM, but are an improvement over the lower-resolution
reanalyses. Explicitly resolved convection and cloud processes (except
microphysics) were not enough to address the model cloud biases.
\end{abstract}

@@

Increasing climate model resolution is one way of improving model accuracy of
representation of the climate system \citeA{mauritsen2022}. It has been
practiced since the advent of climate modelling as more computational power,
memory, and storage capacity become available. It is, however, often not as
easy as changing the grid size because of the complex interplay between model
dynamics and physics, which necessitates adjusting and tuning all components
together.  Increasing resolution is of course limited by the available
computational power and a trade-off with increasing parameterization
complexity, which is another way of improving model accuracy.  Current
computational availability and acceleration from general-purpose computing on
graphics processing units (GPUs) is progressing to enable km-scale (also called
k-scale) Earth system models (ESMs) and coupled atmosphere--ocean general
circulation models (AOGCMs) in research conditions today and operationally in
the forthcoming years.  Therefore, it represents a natural advance in climate
modelling.  Global storm-resolving models (GSRMs) are emerging as a new front
in the development of high-resolution global climate models, with horizontal
grid resolutions of about 2--8 km \citeA{satoh2019,stevens2019}. This is enough
to resolve mesoscale convective storms, but smaller-scale convective plumes and
cloud structure remain unresolved. At an approximately 5-km scale,
non-hydrostatic processes also become important \citeA{weisman1997}, and for
this reason such models are generally non-hydrostatic. The terms global
cloud-resolving models or global convection-permitting/-resolving models are
also sometimes used interchangeably with GSRMs but imply that clouds or
convection are resolved explicitly, which is not entirely true for GSRMs, as
this would require an even higher horizontal resolution \citeA{satoh2019}.
Representative of these efforts is the DYnamics of the Atmospheric general
circulation Modeled On Non-hydrostatic Domains (DYAMOND) project
\citeA{stevens2019,dyamond}, which is an intercomparison of nine global GSRMs
over two 40-day time periods in summer (1 August -- 10 September 2016) and
winter (20 January -- 1 March 2020). A new one-year GSRM intercomparison is
currently proposed by \citeA{takasuka2024}, with the hope of also evaluating the
seasonal cycle and large-scale circulation.  An alternative to using a
computationally costly GSRM is to train an artificial neural network on GSRM
output and use it for subgrid-scale clouds, as done with the GSRM ICON by
\citeA{grundner2022} and \citeA{grundner2023}.

nextGEMS is a project \citeA{nextgems} focused on the research and development
of GSRMs at multiple modelling centres and universities in Europe.  The project
also develops GSRM versions of the Icosahedral Nonhydrostatic Weather and
Climate Model (ICON; \citeA{hohenegger2023}), the Integrated Forecasting System
[IFS; \citeA{ifs2023}], and their ocean components at eddy-resolving
resolutions: ICON-O \citeA{korn2022} coupled with ICON and
Finite-Element/volumE Sea ice-Ocean Model [FESOM; \citeA{wang2014}] and Nucleus
for European Modelling of the Ocean [NEMO; \citeA{madec2023}] coupled with IFS.
The project has so far produced ICON and IFS simulations with three development
versions called Cycle 1--3 and a pre-final version, with a final production
version planned by the end of the project. nextGEMS is not the only project
developing GSRMs; other GSRMs (or GSRM versions of climate models) currently in
development include: Convection-Permitting Simulations With the E3SM Global
Atmosphere Model [SCREAM; \citeA{caldwell2021}], Atmospheric Model [NICAM;
\citeA{satoh2008}], Unified Model (UM), eXperimental System for High-resolution
modeling for Earth-to-Local Domain [X-SHiELD; \citeA{shield}], Action de
Recherche Petite Echelle Grande Echelle-NonHydrostatic version [ARPEGE-NH;
\citeA{bubnova1995,voldoire2017}], Finite-Volume Dynamical Core on the Cubed
Sphere [FV3, \citeA{lin2004}], the National Aeronautics and Space Administration
(NASA) Goddard Earth Observing System global atmospheric model version 5
[GEOS5; \citeA{putman2011}], Model for Prediction Across Scales [MPAS;
\citeA{skamarock2012}], and System for Atmospheric Modeling [SAM;
\citeA{khairoutdinov2003}].

Multiple cloud properties have an effect on shortwave (SW) and longwave (LW)
radiation. To first order, the total cloud fraction, cloud phase, and the
liquid and ice water path are the most important cloud properties influencing
SW and LW radiation. These properties are  in turn influenced by the
atmospheric thermodynamics, convection and circulation, and both the indirect
and direct effects of aerosols. Second order effects on SW and LW radiation are
associated with the cloud droplet size distribution, ice crystal habit, cloud
lifetime, and direct radiative interaction with aerosols.  In the
6$^\mathrm{th}$ phase of the Coupled Model Intercomparison Project [CMIP6;
\citeA{eyring2016}], the cloud feedback has increased relative to CMIP5
\citeA{zelinka2020}, which is one of the main reasons for the higher climate
sensitivity of CMIP6 models.

The Southern Ocean (SO) is known to be a problematic region for climate model
biases \citeA{schuddeboom2021,hyder2018,cesana2022,zhao2022} due to a lack of
surface and in situ observations and being a lower priority region for
numerical weather prediction (NWP) and climate model development because of its
distance from populated areas.  Nevertheless, radiation biases and changes over
an area of its size have a substantial influence on the global climate
\citeA{rintoul2011}, such as affecting the Earth radiation balance, ocean heat
and carbon uptake \citeA{williams2023}, and the SO is an important part of the
global ocean conveyor belt \citeA{wang2014}. In general, marine clouds have a
disproportionate effect on top of atmosphere (TOA) SW radiation due to the
relatively low albedo of the sea surface.  The relative longitudinal symmetry
of the SO means that model cloud biases tend to be similar across longitudes.

Here, we conventionally refer to the SO as ocean regions south of 40°S,
low-latitude SO as 40--55°S and high-latitude SO as south of 55°S. The reason
for this dividing latitude is to split the SO into about two equal zones, as
well as the results by \citeA{schuddeboom2021} (Fig. 2b), showing a contrast in
CMIP model radiation biases, \citeA{schuddeboom2019} (Fig. 2) and
\citeA{kuma2020} (Fig. 3) showing contrasting radiation biases in the Hadley
Centre Global Environmental Model, and \citeA{cesana2022} showing contrasting
cloud biases due to the 0°C isotherm reaching the surface at 55°S. The findings
of \citeA{niu2024}, however, support a different dividing line of 62°S based on
cloud condensation nuclei concentration.

SO radiation biases have been relatively large and systematic compared to the
rest of the globe since at least CMIP3 \citeA{trenberth2010}, and the SO SW
cloud radiative effect (CRE) bias is still positive in eight analysed CMIP6
models analysed by \citeA{schuddeboom2021} over the high-latitude SO, whereas
over the low-latitude SO it tends to be more neutral or negative in some
models. Too much absorbed SW radiation over the SO was also identified in the
GSRM SCREAM \citeA{caldwell2021}. Compensating biases are possible, such as the
`too few too bright' cloud bias, characterised by too small cloud fraction and
too large cloud albedo \citeA{wall2017,kuma2020}, previously described by
\citeA{webb2001}, \citeA{weare2004}, \citeA{zhang2005}, \citeA{karlsson2008},
\citeA{nam2012}, \citeA{klein2013}, and \citeA{bender2017} in other regions and
models, which means that a model can maintain a reasonable SW radiation balance
by reflecting too much SW radiation from clouds, but these cover too small
area.  A study by \citeA{konsta2022} showed that this type of bias is still
present in six analysed CMIP6 models in tropical marine clouds, using the
GCM-Oriented Cloud-Aerosol Lidar and Infrared Pathfinder Satellite Observation
(CALIPSO) Cloud Product [CALIPSO--GOCCP; \citeA{chepfer2010}] and Polarization
\& Anisotropy of Reflectances for Atmospheric Sciences coupled with
Observations from a Lidar [PARASOL; \citeA{lier2008}] as a reference. They
suggest improper simulation of subgrid-scale cloud heterogeneity as a cause.
Compensating cloud biases in the Australian Community Climate and Earth System
Simulator (ACCESS) – Atmosphere-only model version 2 (AM2) over the SO were
analysed by \citeA{fiddes2022} and \citeA{fiddes2024}.  \citeA{possner2022} showed
that over the SO, the DYAMOND GSRM ICON underestimates low-level cloud fraction
on the order of 30\% [relative to Moderate Resolution Imaging Spectroradiometer
(MODIS) data] and overestimates downwelling TOA SW radiation by approximately
10 Wm$^\mathrm{-2}$ [relative to the Clouds and the Earth’s Radiant Energy
System (CERES) data] in the highest model resolution run (2.5 km).
\citeA{zhao2022} reported a similar SW radiation bias in five analysed CMIP6
models over high-latitude SO and the total cloud fraction underestimation on
the order of 10\% over the entire 40--60°S SO. Recently, \citeA{ramadoss2024}
analysed 48 hours of km-scale ICON limited area model NWP simulations over a SO
region adjacent to Tasmania against the Clouds, Aerosols, Precipitation,
Radiation, and atmospherIc Composition Over the southeRn oceaN (CAPRICORN)
voyage cloud and precipitation observations \citeA{mcfarquhar2021}. They found
the ICON cloud optical thickness was underestimated relative to Himawari‐8
satellite observations, but also identified large differences in cloud top
phase.

In general, sea surface temperature (SST) biases in the SO can originate either
in the atmosphere \citeA{hyder2018}, caused by too much shortwave heating of
the surface, too little longwave cooling of the surface, or in the ocean
circulation.  Interactions of both are also possible, for example SST affecting
clouds and clouds affecting the surface radiation. Using the European Centre
for Medium-Range Weather Forecasts (ECMWF) Reanalysis 5 (ERA5) as a reference,
\citeA{zhang2023} have shown that SST biases have improved in CMIP6 compared to
CMIP5, with SST overall increasing in the later CMIP phase. However, over the
SO this resulted in an even higher positive bias, especially in the Atlantic
Ocean (AO) sector of the SO, increasing by up to 1°C.  \citeA{luo2023}
identified that the SO SST bias in an ensemble of 18 CMIP6 models originates
not from the surface heat and radiation fluxes (using reanalyses as a
reference), but from a warm bias in the Northern Atlantic Deep Water.
